Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-06-01 manuscript.Rmd
Edit and shift Social Movements and Civil Wars section

@@

#### Cohesion in Social Movements and Civil Wars
People join social and political resistance movements because they feel a sense
of connection to the cause, and because they have been sufficiently convinced to
participate. Given that people are generally risk averse, overcoming inertia
requires a strong connection to the group identity, or against an individual,
event, or political injustice [@Goldstone_1980; Goldstone_2011]. Political
resistance offers participants the opportunity to “lose themselves” and
relinquish their individual identity in favor of identification with a larger
group identity, thus contributing to a cohesive group ready and mobilized for
action [@Hoffer_2011. Previous research on social movements supports the idea
that participants are likely to join movements when they know and have a
personal connection to someone else already involved [@McAdam_1986. The type of
collective contentious behavior that manifests in society is a function of the
available political opportunity structures. The political opportunity structure,
as defined by Kitschelt (1986) are “comprised of specific configuration of
resources, institutional arrangements and historical precedents for social
mobilization, which facilitate the development of protest movements in some
instances and constrain them in others [@Kitschelt_1986].” What emerges as
strategies and tactics in contentious politics is in essence, a function of what
the system makes possible. This includes the number of people and their
disposition, i.e. nonviolent or violent, as well as their internal structure.
The more cohesive a group is, the greater threat they pose to the establishment.

Cohesion is also a familiar concept in the study of civil conflicts as it
pertains to group affiliation and the ability to overcome collective action
problems. Olson’s (Olson, 1965) seminal work on collective action informs the
theory of group behavior. Individuals must act in concert – cohesively – to
accomplish the task of challenging the authority of the state. Building on
Eisinger’s (1973) analysis of social protest movements [@Eisinger_1973, McAdam,
Tilly and Tarrow [@Tilly_2004, and Francisco [@Francisco_1995 have provided
foundational work on the dynamics of social movements and contentious behavior.
More recently, Gates (2002) considers rebel recruitment for participation in
civil conflicts as a function of internal cohesion [@Gates_2002]. He notes that
rebel movement success is influenced by the degree of group cohesion, which
itself is a function of geographic distance, ideology and ethnicity.

There are multiple rebel groups within Syria, some with fighters sourced locally
from the disaffected Syrian population, and an increasing number of foreign
fighters from transnational terrorist organizations. Gates’ work helps us
understand the intergroup competition for participation in the rebel movement,
as well as the groups’ abilities to punish defectors for nonparticipation.
Foreign fighters would appear to be amply motivated to fight, given their
decision to relocate from their home country which is sometimes in the region
but other times requires them to travel large distances.

Similarly, Sambanis and Shayo (2013) find in their model that “intense ethnic
conflict makes people care more about their ethnic group relative to other
groups, and seek to resemble it more [@Sambanis_Shayo_2013].” As ethnic conflict
intensifies, groups seeking to become more quintessentially like the typical
member of their ethnic group, this should be manifested in their language on
social media. Social medial should reflect this delineation amongst ethnic
groups and between them. However, this may also include a general identification
across ethnic groups in a membership of an opposition to an oppressive leader.

It is this general identification with the larger opposition movement, not the
individual identity groups within Syria, which we seek to understand in this
paper. It is implausible to expect there to be one cohesive narrative from the
multiple groups with disparate incentives. However, they do have important
similarities, namely, they are all interested in opposing the current Syrian
regime using largely violent means, although there are notable and regular
instances of nonviolent protests on Fridays, following prayer services. Thus,
what we are able to capture in this study is the degree to which all opposition
movements to the Syrian government have a similar content and tone, with a
similar ebb and flow. It should be increasingly threatening to the regime if the
rebel movement across the disparate groups is growing stronger and more unified,
represented by similar language used across groups.

The civil war literature has further explored the role of social cohesion in
detail. Collier and Hoeffler identified the role of social homogeneity as having
an increased likelihood on social conflict, especially when the society is
polarized [@Collier_Hoeffler_2004]. Cederman and Girardin  write that, “In
general, nationalist groups derive their cohesion from collective-level
identities that resonate deeply with the emotional needs of modern citizens.”
Sambanis and Schulhofer-Wohl note that “group cohesion and solidarity (underlie)
the offensive advantage (Sambanis & Schulhofer-Wohl, 2009).” Anyanwu (2002)
writes, “Another source of rebel military opportunity is social cohesion…an
ethnically diverse society may reduce the opportunity for rebellion by limiting
the recruitment pool [@Anyanwu_2002].” Referencing Gurr’s work on minorities at
risk, Regan and Norton (2005) note that “a religious or linguistic minority
might suffer disproportionately in a given society, and this form of
aggrievement can lead to unrest across the social lines that distinguish the
minority group. Therefore, while collective cohesion is a function of targeted
maldistributions, it is individual penury that allows for this collective
mobilization [@Gurr_2000; @Regan_Norton_2005].” Thus scholarship concurs that
group cohesion matters to group mobilization; we add to this concept by
introducing the role of language in facilitating cohesion.

## Sociolinguistic cohesion: Connectedness of speakers, events, and ideas
We tie the constructs of lexical, interlocutor, and social/physical cohesion
together in a theoretical framework called sociolinguistic cohesion, as shown in
Figure \@ref(fig:theoretical-framework). We can measure the changes in cohesion
preceding and in response to external stimuli by evaluating the messages and the
context in which they occur. Several mechanisms are in motion here. Previous
values of linguistic cohesion should influence future values of linguistic
cohesion. They should also influence future actions in the real world space.
Previous events influence future events, and also the future cohesiveness of the
opposition.

>>>>

## The Role of Cohesion in Social Movements and Civil Wars

People join social and political resistance movements because they feel a sense
of connection to the cause and because they have been sufficiently convinced to
participate. Given that people are generally risk-averse, overcoming inertia
requires a strong connection to the group identity or a strong aversion to an individual,
event, or political injustice [@goldstone1980theories; @goldstone2011understanding]. 
Political resistance offers participants the opportunity to “lose themselves” and
relinquish their individual identity in favor of identification with a larger
group identity, contributing to a cohesive group ready and mobilized for
action [@Hoffer2011true].

Previous research on social movements supports the idea that participants are
likely to join movements when they know and have a personal connection to
someone else already involved [@mcadam1986recruitment]. The type of collective
contentious behavior that manifests in society is a function of the available
political opportunity structures. The political opportunity structure are
“comprised *(sic)* of specific configuration of resources, institutional
arrangements and historical precedents for social mobilization, which facilitate
the development of protest movements in some instances and constrain them in
others” [@Kitschelt_1986]. What emerges as strategies and tactics in contentious
politics are, in essence, a function of what the system makes possible. This
includes the number of people and their disposition (i.e. nonviolent or
violent), as well as their internal structure. The more cohesive a group is, the
greater threat they pose to the establishment.

Cohesion is also a familiar concept in the study of civil conflicts, as it
pertains to group affiliation and the ability to overcome collective action
problems. Individuals must act in concert---that is, cohesively---to
accomplish the task of challenging the authority of the state [@olson1965logic;
@eisinger1973conditions; @tilly2004contentious; @francisco1995relationship].
Recent work has considered rebel recruitment for participation in
civil conflicts as a function of internal cohesion [@gates2002recruitment]:
Rebel movement success is influenced by the degree of group cohesion, which
itself is a function of geographic distance, ideology, and ethnicity.
Within Syria, multiple rebel groups within Syria exist, sourcing fighters locally
from the disaffected Syrian population and internationally through
fighters from transnational terrorist organizations. There is considerable 
intergroup competition for participation in the rebel movement and
punishment for defectors for nonparticipation [@gates2002recruitment].
Foreign fighters would appear to be amply motivated to fight, given their
decision to relocate from a home country that may be located in the region
or much farther away.

Another factor influencing social cohesion during social movements and
civil wars is ethnic identity and the degree of ethnic conflict associated with
the movement. Greater degrees of ethnic conflict increase the salience of
ethnic identity, leading people to identify more strongly with their own
ethnic identity [@sambanis2009s]. As ethnic conflict
intensifies, groups seek to become more quintessentially like the typical
member of their ethnic group. This can manifest in a number of ways, including
in behavior on social media. However, this may also include a general identification
across ethnic groups in a membership of an opposition to an oppressive leader.

The civil war literature has further explored the role of social cohesion in
detail. Social homogeneity increases the likelihood of social conflict,
especially when the society is polarized [@collier2004greed]; this is partly
due to nationalist groups' strong identification with a collective identity
[@cederman2007beyond] and their social solidarity [@sambanis2009s]. These
same processes can be similarly important for rebels [@anyanwu2002economic].
Referencing Gurr’s [-@gurr2000peoples] work on minorities at
risk, Regan and Norton [-@regan2005greed] note that

>a religious or linguistic minority
might suffer disproportionately in a given society, and this form of
aggrievement can lead to unrest across the social lines that distinguish the
minority group. Therefore, while collective cohesion is a function of targeted
maldistributions, it is individual penury that allows for this collective
mobilization. [@regan2005greed]

The body of previous work concurs that group cohesion matters to group
mobilization. We build on this by providing a quantitative approach to
understanding cohesion from multiple levels and by incorporating analytic tools
that allow us to investigate the interconnectedness of those levels.

@@

### Sociolinguistic cohesion: Connectedness of speakers, events, and ideas
We tie the constructs of lexical, interlocutor, and social cohesion
together in a theoretical framework called sociolinguistic cohesion, as shown in
Figure \@ref(fig:theoretical-framework). We can measure the changes in cohesion
preceding and in response to external stimuli by evaluating the messages and the
context in which they occur. Several mechanisms are in motion here. Previous
values of linguistic cohesion should influence future values of linguistic
cohesion. They should also influence future actions in the real world space.
Previous events influence future events, and also the future cohesiveness of the
opposition.
